---
output:
  html_document: 
    self_contained: yes
    toc: true
    toc_float: true
    toc_collapsed: true
    number_sections: true
    theme: cosmo
params:
  phenotype_name: "phenotype of interest"
  color_code_yaml: NA
  input_prefix: NA
  rlib: NA
  ntop: 10
  bxcan_pval_col: 'pval'
  fdr_cutoff: 0.05
title: "`r paste0('BrainXcan results for ', params$phenotype_name)`"
---


```{r setup, include=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(pander))
options(stringsAsFactors = FALSE)
theme_set(theme_classic(base_size = 12))
panderOptions('table.split.table', Inf)
source(params$rlib)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

```{r parameters}
input_prefix = params$input_prefix
meta_file = params$meta_file
color_code_yaml = params$color_code_yaml
phenotype_name = params$phenotype_name
ntop = params$ntop
bxcan_pval_col = params$bxcan_pval_col
fdr_cutoff = params$fdr_cutoff
pval_col_to_name = list(
  pval = 'raw p-values from BrainXcan',
  pval_adj_emp_null = 'adjusted p-values based on simulated weights',
  pval_adj_perm_null = 'adjusted p-values based on permuted weights',
  pval_adj_vc = 'variance control adjusted p-values'
)
pval_col_to_zscore_col = list(
  pval = 'z_brainxcan',
  pval_adj_emp_null = 'z_adj_perm_null',
  pval_adj_perm_null = 'z_adj_perm_null',
  pval_adj_vc = 'z_adj_vc'
)
```

```{r load_meta}
# gannot = read.csv(meta_file)
color_map = load_color_code_yaml(color_code_yaml)
```

```{r load_results}
bxcan_file = paste0(input_prefix, '.sbrainxcan.csv')
sbxcan = read.csv(bxcan_file)
z_col = pval_col_to_zscore_col[[bxcan_pval_col]]
if(!is.null(z_col)) {
  if(z_col %in% colnames(sbxcan)) {
    sbxcan$zscore = sbxcan[[z_col]]
  } else {
    sbxcan = sbxcan %>% mutate(zscore = abs( qnorm(!!sym(bxcan_pval_col)/2) ) * sign(bhat))
  }
} else {
  sbxcan = sbxcan %>% mutate(zscore = abs( qnorm(!!sym(bxcan_pval_col)/2) ) * sign(bhat))
}
allxcan = sbxcan
```

```{r subset_to_idp_main_results}
allxcan_annot = allxcan # %>% left_join(gannot %>% select(-t1_or_dmri), by = "IDP") 
  
## PC's should have region==PC
ind = substr(allxcan_annot$IDP, 1, 2) == "PC"
allxcan_annot$region[ind] = allxcan$IDP[ind]
## set subtype of PCs as subtype
ind = substr(allxcan_annot$IDP, 1, 2) == "PC"
allxcan_annot$subtype[ind] = allxcan$IDP[ind]

## limit to IDP main results: T1 and TBSS-based dMRI
xcandf = allxcan_annot %>% filter(subtype %in% names(color_map)) %>% rename(side = left_or_right)
```


# BrainXcan results 

In this analysis, we tested `r nrow(allxcan)` structural and diffusion MRI-derived phenotypes with cross validated spearman correlation >= `r signif(min(allxcan$CV_Spearman), digits = 3)`.

<!--- 
```{r, bonforroni_cutoffs}
## define number of total tests per phenotype to be used for bonferroni correction
n_pred_idps = length(unique(xcandf$IDP))
p_thres = 0.05 / n_pred_idps
z_thres = abs(qnorm(p_thres / 2))
``` 
-->

```{r, qvalue}
## define number of total tests per phenotype 
## calculate qvalue
n_pred_idps = length(unique(xcandf$IDP))
xcandf$qvalue = get_qvalue(xcandf[[bxcan_pval_col]])
# if(any(xcandf$qvalue < qval_cutoff)) {
#   p_thres = max(xcandf[[bxcan_pval_col]][xcandf$qvalue < qval_cutoff])
#   z_thres = abs(qnorm(p_thres / 2))
# } else {
#   p_thres = 0
#   z_thres = NULL
# }

``` 


```{r, fdr}
xcandf$fdr = get_fdr(xcandf[[bxcan_pval_col]])
if(any(xcandf$fdr < fdr_cutoff)) {
  p_thres = max(xcandf[[bxcan_pval_col]][xcandf$fdr < fdr_cutoff])
  z_thres = abs(qnorm(p_thres / 2))
} else {
  p_thres = 0
  z_thres = NULL
}

``` 

For the main analysis we focused on `r n_pred_idps` IDPs, which include `r sum(xcandf$subtype == 'Gray-Cortical')` cortical volumes, `r sum(xcandf$subtype == 'Subcortical')` sub-cortical volumes, `r sum(xcandf$subtype == 'Gray-Subcortical')` sub-cortical gray matter volumes.
In below, we show the overall number of IDP by IDP subtype in the main results.

```{r, show_count}
xcandf %>% count(subtype) %>% pander::pander()
```

The p-value being used is `r message(pval_col_to_name[[bxcan_pval_col]])`.

Among `r n_pred_idps` IDPs (`r sum(xcandf$modality=="T1")` T1 and `r sum(xcandf$modality=="dMRI")` diffusion MRI, including the subtype-specific PCs), `r sum(xcandf$fdr < fdr_cutoff)` were significantly associated with `r phenotype_name` (FDR < $`r fdr_cutoff`$). 

 
```{r, vis_t1, fig.height=7, fig.width=8}
idp_type = 'T1'
pp = plot_bxcan_ordered(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
```

```{r, vis_dmri, fig.height=6, fig.width=8}
idp_type = 'dMRI'
pp = plot_bxcan_ordered(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
``` 


<!--- 
```{r, vis_t12, fig.height=14, fig.width=6}
idp_type = 'T1'
pp = plot_bxcan_ordered_heatmap(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
# ggsave(glue::glue('{outdir}/{phenotype_name}.{idp_type}.png'), width = 6, height = 14)
```

```{r, vis_dmri2, fig.height=12, fig.width=6}
idp_type = 'dMRI'
pp = plot_bxcan_ordered_heatmap(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
# ggsave(glue::glue('{outdir}/{phenotype_name}.{idp_type}.png'), width = 6, height = 12)
```
-->


```{r load_mr}
df_mr = load_mr_sum_stats(paste0(input_prefix, '_files/MR_results/'))
if(!is.null(df_mr)) {
  df_mr$mr_pval = gen_mr_sign(df_mr$b, df_mr$pval)
  tmp = df_mr %>% filter(method == 'ACAT meta-analysis') %>% reshape2::dcast(IDP ~ direction, value.var = 'pval')
  xcandf_w_mr = xcandf %>% left_join(tmp, by = 'IDP') 
  xcandf_w_mr = xcandf_w_mr %>% mutate(link = paste0('[MR plot](#', tolower(IDP), ')'))
  xcandf_w_mr$link[ is.na(xcandf_w_mr$`Phenotype -> IDP`) ] = NA
}
```

The top `r ntop` associations for dMRI are 
```{r dmri_top_n, results='asis'}
if(!is.null(df_mr)) {
  xcandf_w_mr %>% select(IDP, modality, subtype, region, side, zscore, fdr, `IDP -> Phenotype`, `Phenotype -> IDP`, link) %>% filter(fdr < fdr_cutoff) %>% filter(modality == "dMRI") %>% arrange(fdr) %>% select(-modality) %>% head(ntop) %>% pander
} else {
  cat('not shown since no significant IDP being identified (pass p-value cutoff).\n\n')
}
```

The top `r ntop` associations for T1 are 
```{r t1_top_n, results='asis'}
if(!is.null(df_mr)) {
  xcandf_w_mr %>% select(IDP, modality, subtype, region, side, zscore, fdr, `IDP -> Phenotype`, `Phenotype -> IDP`, link) %>% filter(fdr < fdr_cutoff) %>% filter(modality == "T1") %>% arrange(fdr) %>% select(-modality) %>% head(ntop) %>% pander
} else {
  cat('not shown since no significant IDP being identified (pass p-value cutoff).\n\n')
}
```

# Visualization of associations in the brain

```{r bxcan_figures, results='asis'}
idp_types = c('T1', 'dMRI')
for(i in idp_types) {
  cat('\n\n')
  cat(paste0('## ', i, '\n\n'))
  fn = paste0(input_prefix, '.vis.', i, '.pdf')
  fn_png = paste0(tempdir(), stringr::str_replace(basename(fn), '\\.pdf', '\\.png'))
  pdftools::pdf_convert(fn, filenames = fn_png, verbose = F, dpi = 300)
  cat(paste0('![](', fn_png, ')'))
  # system(paste0('rm ', fn_png))
}
```

# BrainXcan MR Figures

```{r mr_figures, results='asis'}
if(!is.null(df_mr)) {
  mr_idps = df_mr %>% select(IDP) %>% distinct()
  for(i in mr_idps$IDP) {
    if(i %in% xcandf$IDP) {
      # do next chunks
    } else {
      next
    }
    fig_fn = Sys.glob(paste0(input_prefix, '_files/MR_results/', '*', i, '.MR_vis.pdf'))
    if(length(fig_fn) > 1) {
      next
    }
    info = allxcan_annot %>% filter(IDP == i)
    msg = i
    if(nrow(info) == 1) {
      msg = info$notes
    }
    cat('\n\n')
    cat(paste0('## ', i, ' {#', tolower(i), '}', '\n\n'))
    cat(msg, '\n\n')
    fn_png = paste0(tempdir(), stringr::str_replace(basename(fig_fn), '\\.pdf', '\\.png'))
    pdftools::pdf_convert(fig_fn, filenames = fn_png, verbose = F, dpi = 200)
    cat(paste0('![](', fn_png, ')'))
    # system(paste0('rm ', fn_png))
  }
} else {
  cat('No significant IDP being identified.\n\n')
}
```
